### 2025-05-08 introduction.tex
new

\begin{equation} \label{eq:fnn}
    f(x; \theta) = f^{(k)}(f^{(k-1)} \cdots (f^{(2)}(f^{(1)}(x;\theta^{(1)});\theta^{(2)} )\cdots ;\theta^{(k-1)});\theta^{(k)}),
\end{equation}
where $\theta = (\theta^{(1)}, \theta^{(2)}, \ldots, \theta^{(k)})$ are trainable parameters of the neural network, $f^{(i)}$ is an $i$-th layer of the neural network, $x$ is an input feature vector, $f^{(1)}(x;\theta^{(1)}) = x$, $f^{(k)}$ is an output layer, $k$ is a number of layers, and $f^{(i)}$ is an output of the $i$-th layer.
To transition from layer $i-1$ to layer $i$,
\begin{equation} \label{eq:neural}
    f^{(i)}(z;\theta^{(i)}) = a^{(i)}(\beta_0^{(i)} + \sum_{j=1}^{p_i} \beta_j^{(i)} z),
\end{equation}
where $a^{(i)}$ are nonlinear activation functions, $\theta^{(i)} = (\beta_0^{(i)}, \dots, \beta_p^{(i)})$ are trainable parameters of the $i$-th layer, and $p_i$ is a number of neurons in the previous layer.
Equation~\eqref{eq:neural} represents a single neuron in a neural network, which applies a linear transformation to the input followed by a non-linear activation function.
The linear transformation is similar to the linear regression model in Equation~\eqref{eq:regression}, but the non-linear activation function allows the neural network to model complex, non-linear relationships in the data.
The \gls{fnn} is tuned by adjusting the trainable parameters $\theta$ while fixing the activation functions $a^{(i)}$.
A typical choice for the activation function is the \gls{relu}, which is defined as $a(z) = \max(0, z)$~\citep{nair2010rectified}.
The activation function choice is crucial for training neural networks.
While sigmoid and hyperbolic tangent functions were popular in early neural networks, ReLU has become the default activation function for deep networks due to its ability to mitigate the vanishing gradient problem and promote sparse activations~\citep{lecun2015deep}.

>>>>

\begin{equation} \label{eq:fnn}
    f(x; \theta) = f^{(k)}(f^{(k-1)} \cdots (f^{(2)}(f^{(1)}(x;\theta^{(1)});\theta^{(2)} )\cdots ;\theta^{(k-1)});\theta^{(k)}),
\end{equation}
where $\theta = (\theta^{(1)}, \theta^{(2)}, \ldots, \theta^{(k)})$ are trainable parameters of the neural network, $f^{(i)}$ is an $i$-th layer of the neural network, $x$ is an input feature vector, $f^{(1)}(x;\theta^{(1)}) = x$, $f^{(k)}$ is an output layer, $k$ is a number of layers, and $f^{(i)}$ is an output of the $i$-th layer.
To transition from layer $i-1$ to layer $i$,
\begin{equation} \label{eq:neural}
    f^{(i)}(z;\theta^{(i)}) = a^{(i)}(\beta_0^{(i)} + \sum_{j=1}^{p_i} \beta_j^{(i)} z),
\end{equation}
where $a^{(i)}$ are nonlinear activation functions, $\theta^{(i)} = (\beta_0^{(i)}, \dots, \beta_p^{(i)})$ are trainable parameters of the $i$-th layer, and $p_i$ is a number of neurons in the previous layer.
Equation~\eqref{eq:neural} represents a single neuron in a neural network, which applies a linear transformation to the input followed by a non-linear activation function.
The linear transformation is similar to the linear regression model in Equation~\eqref{eq:regression}, but the non-linear activation function allows the neural network to model complex, non-linear relationships in the data.
The \gls{fnn} is tuned by adjusting the trainable parameters $\theta$ while fixing the activation functions $a^{(i)}$.
A typical choice for the activation function is the \gls{relu}, which is defined as $a(z) = \max(0, z)$~\citep{nair2010rectified}.
The activation function choice is crucial for training neural networks.
While sigmoid and hyperbolic tangent functions were popular in early neural networks, \gls{relu} has become the default activation function for deep networks due to its ability to mitigate the vanishing gradient problem and promote sparse activations~\citep{lecun2015deep}.

### 2025-05-08 project1/nsSurvey.tex
new

\section{Finite-Sample Experiments}\label{sec1:numerical-experiments}
The theoretical framework has lent itself for a comparison of the asymptotic convergence behavior for different nested simulation procedures.
However, a simulation budget in practice is almost always finite.
In this section, we conduct a series of numerical experiments to compare the empirical convergence of the nested simulation procedures for different risk measures, option types, and asset dimensions.
Numerical experiments are conducted on portfolios that consist of options on $d$ underlying assets, whose dynamics follow a multidimensional geometric Brownian motion with $0.3$ pairwise correlation.
The initial values of the assets are set to be $100$.
The drifts are set to be $0.08$, and the risk free rate is set to be $0.05$.

@@

Our previous numerical experiments have been conducted under the assumption that the underlying asset dynamics follow a multidimensional geometric Brownian motion with $0.3$ pairwise correlation.
To examine the sensitivity of the empirical convergence behavior to the asset model, we consider a stochastic volatility model of a Heston type~\citep{heston1993closed}.
In a Heston model, the asset price and the volatility are correlated, and the volatility is a stochastic process.

@@

In Figure~\ref{fig1:sens_model}, the empirical convergence results of the regression-based nested simulation procedure for a quadratic tracking error and a 90\%-\gls{cvar} are illustrated for different asset models, i.e., a geometric Brownian motion and a Heston model.
The empirical rates of convergence of the regression-based nested simulation procedure are observed to be insensitive to the asset model.

>>>>

\section{Finite-Sample Experiments}\label{sec1:numerical-experiments}
The theoretical framework has lent itself for a comparison of the asymptotic convergence behavior for different nested simulation procedures.
However, a simulation budget in practice is almost always finite.
In this section, we conduct a series of numerical experiments to compare the empirical convergence of the nested simulation procedures for different risk measures, option types, and asset dimensions.
Numerical experiments are conducted on portfolios that consist of options on $d$ underlying assets, whose dynamics follow a multidimensional \gls{gbm} with $0.3$ pairwise correlation.
The initial values of the assets are set to be $100$.
The drifts are set to be $0.08$, and the risk free rate is set to be $0.05$.

@@

Our previous numerical experiments have been conducted under the assumption that the underlying asset dynamics follow a multidimensional \gls{gbm} with $0.3$ pairwise correlation.
To examine the sensitivity of the empirical convergence behavior to the asset model, we consider a stochastic volatility model of a Heston type~\citep{heston1993closed}.
In a Heston model, the asset price and the volatility are correlated, and the volatility is a stochastic process.

@@

In Figure~\ref{fig1:sens_model}, the empirical convergence results of the regression-based nested simulation procedure for a quadratic tracking error and a 90\%-\gls{cvar} are illustrated for different asset models, i.e., a \gls{gbm} and a Heston model.
The empirical rates of convergence of the regression-based nested simulation procedure are observed to be insensitive to the asset model.

### 2025-05-08 project2/nnMetamodels.tex
new

The risk-free rate is $0.2\%$ per month and the underlying asset $S_t$ is modeled by a regime-switching geometric Brownian motion with parameters specified in Table~\ref{tab:regime_params}.

\begin{table}[ht]
    \centering
    \begin{tabular}{lc}
        \toprule
        \textbf{Parameter}  & \textbf{Monthly Value} \\
        \midrule
        Risk-free rate & $0.002$ \\
        Mean return for regime 1  & $0.0085$ \\
        Mean return for regime 2   & $-0.0200$ \\
        Standard deviation for regime 1  & $0.035$ \\
        Standard deviation for regime 2  & $0.080$ \\
        Transition probability from regime 1  & $0.04$ \\
        Transition probability from regime 2  & $0.20$ \\
        \bottomrule
    \end{tabular}
    \caption{Real-world parameters for the regime-switching model (monthly rates)}
    \label{tab:regime_params}
\end{table}

@@

Five metamodels are considered in this experiment: \gls{mlr}, \gls{qpr} without interaction terms, \gls{fnn}, \gls{rnn}, and \gls{lstm} network.
\gls{mlr} and \gls{qpr} are two common metamodels in the simulation literature, and their implemention is straightfoward in our simulation setting.
\gls{fnn} (Equation~\refeq{eq:fnn}) is a generic neural network while \gls{rnn} (Equation~\refeq{eq:rnn}) and \gls{lstm} (Equation~\refeq{eq:lstm}) are specialized models to accommodate the sequential structure of our time-series features.
A $\tanh$ activation function is used for \gls{rnn} and \gls{lstm} layers, and a ReLU activation function is used for the fully-connected layers.
All neural network metamodels are trained by the Adam optimizer~\citep{kingma2014adam} with an initial learning rate of $0.001$ and an exponential learning rate decay schedule.
All \gls{dnn} metamodels are trained with a dropout rate of $10\%$.
The architectures and training settings are typical choices in the deep learning literature.
The training labels are normalized to have a zero mean and a unit standard deviation.
The numbers of trainable parameters are recorded in the second column of Table~\ref{tab:gmwb_arch}.
We observe that the three \gls{dnn} metamodels have orders of magnitudes more trainable parameters than the two traditional regression metamodels.
Following the convention in machine learning literature, the three \gls{dnn} metamodels are said to have much higher \textit{model capacities}.
Treating regression metamodels as shallow neural networks, \gls{mlr} and \gls{qpr} are two neural networks with no hidden layer. 
For a traditional regression metamodel, the numbers of neurons in its input layer equals to the cardinality of its regression basis.
On the other hand, the three \gls{dnn} metamodels have several hidden layers and are more flexible than \gls{mlr} and \gls{qpr}.
Our \gls{fnn} has two hidden layers prior to its output layer.
Its hidden layers are feature extraction layers that transform the 240-dimension feature vector into a 32-dimension vector that can directly relate to the target variable.
This extraction process is automatic and does not require manual feature engineering.
Our \gls{rnn} and \gls{lstm} networks have three hidden layers.
They contain similar numbers of neurons as the \gls{fnn} but have deeper network structures.
For \gls{rnn}, two of its hidden layers are recurrent layers that capture the temporal dependency in the feature, and its last hidden layer is a standard feedforward layer that transforms the previous hidden states into a 32-dimension vector for target prediction.
Our \gls{lstm} has the same architecture as the \gls{rnn} but with \gls{lstm} cells in place of \gls{rnn} cells.
.

>>>>

The risk-free rate is $0.2\%$ per month and the underlying asset $S_t$ is modeled by a \gls{rsgbm} with parameters specified in Table~\ref{tab:regime_params}.

\begin{table}[ht]
    \centering
    \begin{tabular}{lc}
        \toprule
        \textbf{Parameter}  & \textbf{Monthly Value} \\
        \midrule
        Risk-free rate & $0.002$ \\
        Mean return for regime 1  & $0.0085$ \\
        Mean return for regime 2   & $-0.0200$ \\
        Standard deviation for regime 1  & $0.035$ \\
        Standard deviation for regime 2  & $0.080$ \\
        Transition probability from regime 1  & $0.04$ \\
        Transition probability from regime 2  & $0.20$ \\
        \bottomrule
    \end{tabular}
    \caption{Real-world parameters for the \gls{rsgbm} model (monthly rates)}
    \label{tab:regime_params}
\end{table}

@@

Five metamodels are considered in this experiment: \gls{mlr}, \gls{qpr} without interaction terms, \gls{fnn}, \gls{rnn}, and \gls{lstm} network.
\gls{mlr} and \gls{qpr} are two common metamodels in the simulation literature, and their implemention is straightfoward in our simulation setting.
\gls{fnn} (Equation~\refeq{eq:fnn}) is a generic neural network while \gls{rnn} (Equation~\refeq{eq:rnn}) and \gls{lstm} (Equation~\refeq{eq:lstm}) are specialized models to accommodate the sequential structure of our time-series features.
A $\tanh$ activation function is used for \gls{rnn} and \gls{lstm} layers, and a \gls{relu} activation function is used for the fully-connected layers.
All neural network metamodels are trained by the Adam optimizer~\citep{kingma2014adam} with an initial learning rate of $0.001$ and an exponential learning rate decay schedule.
All \gls{dnn} metamodels are trained with a dropout rate of $10\%$.
The architectures and training settings are typical choices in the deep learning literature.
The training labels are normalized to have a zero mean and a unit standard deviation.
The numbers of trainable parameters are recorded in the second column of Table~\ref{tab:gmwb_arch}.
We observe that the three \gls{dnn} metamodels have orders of magnitudes more trainable parameters than the two traditional regression metamodels.
Following the convention in machine learning literature, the three \gls{dnn} metamodels are said to have much higher \textit{model capacities}.
Treating regression metamodels as shallow neural networks, \gls{mlr} and \gls{qpr} are two neural networks with no hidden layer. 
For a traditional regression metamodel, the numbers of neurons in its input layer equals to the cardinality of its regression basis.
On the other hand, the three \gls{dnn} metamodels have several hidden layers and are more flexible than \gls{mlr} and \gls{qpr}.
Our \gls{fnn} has two hidden layers prior to its output layer.
Its hidden layers are feature extraction layers that transform the 240-dimension feature vector into a 32-dimension vector that can directly relate to the target variable.
This extraction process is automatic and does not require manual feature engineering.
Our \gls{rnn} and \gls{lstm} networks have three hidden layers.
They contain similar numbers of neurons as the \gls{fnn} but have deeper network structures.
For \gls{rnn}, two of its hidden layers are recurrent layers that capture the temporal dependency in the feature, and its last hidden layer is a standard feedforward layer that transforms the previous hidden states into a 32-dimension vector for target prediction.
Our \gls{lstm} has the same architecture as the \gls{rnn} but with \gls{lstm} cells in place of \gls{rnn} cells.
.
